---
title: "centrality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Centrality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In network analysis, centrality measures help identify the most important vertices within a graph. These measures are used in a wide variety of domains such as sociology (influential individuals), biology (essential proteins), or web analysis (important web pages).

The `igraph` package implements the following indices:

- degree (`degree()`)
- weighted degree (`strength()`)
- betweenness (`betweenness()`)
- closeness (`closeness()`)
- eigenvector (`eigen_centrality()`)
- alpha centrality (`alpha_centrality()`)
- power centrality (`power_centrality()`)
- PageRank (`page_rank()`)
- eccentricity (`eccentricity()`)
- hubs and authorities (`authority_score()` and `hub_score()`)
- subgraph centrality (`subgraph_centrality()`)



```{r setup}
library(igraph)
```

The following graph will be used to illustrate the undirected centrality measures.
```{r graph}
el <- matrix(
  c(
    "A", "B", "C", "C", "D", "E", "E", "F", "F", "F", "F", "G", "G", "G", "H",
    "H", "I", "K", "D", "E", "K", "H", "I", "K", "G", "H", "J", "K", "I", "J",
    "K", "I", "J", "J"
  ),
  nrow = 17L,
  ncol = 2L
)

g <- graph_from_edgelist(el, directed = FALSE)
plot(
  g,
  vertex.label = V(g)$name,
  vertex.size = 20
)
```

## Undirected and unweighted centrality indices

### Degree centrality

The degree centrality of a vertex is the number of edges incident to it. In an undirected graph, this is simply the number of neighbors.

```{r degree}
degree(g)
```

### Betweenness centrality

Betweenness centrality measures the extent to which a vertex lies on the shortest paths between other vertices. It is calculated as the sum of the fraction of all-pairs shortest paths that pass through the vertex.

```{r betweenness}
betweenness(g)
```

### Closeness centrality
Closeness centrality is the reciprocal of the average length of the shortest path from a vertex to all other vertices in the graph. It indicates how quickly a vertex can access other vertices.

```{r closeness}
closeness(g)
```

### Eigenvector centrality

Eigenvector centrality measures a vertex's influence based on the number and quality of its connections. It considers not only the number of neighbors but also the centrality of those neighbors.

```{r eigenvector}
eigen_centrality(g)$vector
```

### Subgraph centrality

Subgraph centrality is a measure that counts the number of closed walks of all lengths that pass through a vertex. It provides a more comprehensive view of a vertex's importance in the graph structure.

```{r subgraph_centrality}
subgraph_centrality(g)
```

## Undirected and weighted centrality measures

When the graph is weighted, the centrality measures can be adjusted to account for edge weights. The `strength()` function calculates the weighted degree, and other measures can also be adapted to consider weights.
