---
title: "Generators for Graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph-generators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(igraph)
```

The `igraph` package provides a variety of functions to generate different types of graphs. This vignette will cover some of the most commonly used graph generators.

## Trivial graph generators

By trivial, we mean graphs that are not generated from any particular model or structure, but rather are created with specific parameters. These include:
- **Empty graph**: A graph with no vertices or edges.
- **Complete graph**: A graph where every pair of distinct vertices is connected by a unique edge.
- **Star graph**: A graph with one central vertex connected to all other vertices.
- **Path graph**: A simple path with a specified number of vertices.
- **Cycle graph**: A closed path where the last vertex is connected back to the first.
- **Lattice graph**: A grid-like structure where vertices are arranged in a regular pattern.
- **Tree graph**: A connected acyclic graph with a specified number of vertices.

The code below shows how to create these trivial graphs using the `igraph` package. The examples use undirected graphs, but you can set `directed = TRUE` for directed versions or change the `mode` parameter as needed. In all cases, the parameter `n` specifies the number of vertices in the graph.

```{r example_trivial_graphs}
g1 <- make_empty_graph(n = 10, directed = FALSE)
plot(g1)
g2 <- make_full_graph(n = 10, directed = FALSE)
plot(g2)
g3 <- make_star(n = 10, mode = "undirected")
plot(g3)
g4 <- make_ring(n = 10, directed = FALSE, circular = FALSE)
plot(g4)
g5 <- make_ring(n = 10, directed = FALSE, circular = TRUE)
plot(g5)
g6 <- make_lattice(dimvector = c(3, 3), directed = FALSE)
plot(g6)
```

## Graph generators from models
The `igraph` package also includes generators based on specific models, such as:
- **Erdős-Rényi model**: Generates random graphs where each edge is included with a certain probability.
- **Barabási-Albert model**: Generates scale-free networks using a preferential attachment mechanism.
- **Watts-Strogatz model**: Generates small-world networks with a specified degree of randomness.
- **Configuration model**: Generates random graphs with a specified degree sequence.


```{r sample_gnp}
g7 <- sample_gnp(n = 100, p = 0.05, directed = FALSE)
g7
```

The parameters `n` and `p` control the number of vertices and the probability of edge creation, respectively. The resulting graph is a random graph where each edge is included with probability `p`.

```{r sample_pa}
g8 <- sample_pa(n = 100, power = 1, directed = FALSE)
g8
```

The `power` parameter controls the preferential attachment mechanism, which influences the degree distribution of the resulting graph.

```{r sample_ws}
g9 <- sample_smallworld(dim = 1, size = 100, nei = 2, p = 0.05, directed = FALSE)
g9
```
The `dim` parameter specifies the dimensionality of the small-world network, `size` is the number of vertices, `nei` is the number of nearest neighbors to connect to, and `p` is the probability of rewiring edges.
