---
title: "Improving igraph interface: current deprecations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


To provide a more consistent interface, part of igraph functions or arguments are slowly but consistently being deprecated.
You can follow along the current level of deprecations in the table below.

Please update your codebases as soon as you can, be they packages or scripts.
To do that, run your code in a session where you installed the latest igraph version, possibly even the development version that you can install from R-universe:

```r
options(
  repos = c(
    igraph = 'https://igraph.r-universe.dev',
    CRAN = 'https://cloud.r-project.org'
  )
)
install.packages('igraph')
```

And pay attention to any message emitted through the lifecycle package.


Thank you for your cooperation!
Direct any question to us in the [igraph issue tracker](https://github.com/igraph/rigraph/issues).


```{r, echo=FALSE}
clean_arg <- function(arg) {
  arg <- sub("^'", "", arg)
  arg <- sub("'$", "", arg)
  arg <- sub('^"', '', arg)
  arg <- sub('"$', '', arg)
  arg
}

link_new <- function(new) {
  # downlit can't link something like "igraph::assortativity(values =)"
  new_str <- sub("\\(.*", "", new)
  url <- downlit::autolink_url(sprintf("igraph::%s", new_str))
  if (is.na(url)) {
    new
  } else {
    sprintf('<a href="%s">%s</a>', url, new)
  }
}

parse_lifecycle_call <- function(symbol_function_call) {
  level <- sub("deprecate_", "", xml2::xml_text(symbol_function_call))
  level <- switch(
    level,
    soft = "1 -- soft",
    warn = "2 -- warn",
    stop = "3 -- stop"
  )
  
  args <- purrr::keep(
    xml2::xml_siblings(xml2::xml_parent(symbol_function_call)),
    \(x) xml2::xml_name(x) == "expr"
  )
  version <- xml2::xml_text(args[[1]])
  old <- xml2::xml_text(args[[2]])
  new <- if (length(args) > 2) xml2::xml_text(args[[3]]) else ""
  
  new_text <- link_new(clean_arg(new))
  
  if (!grepl("<a", new_text)) {
    old_text <- link_new(clean_arg(old))
  } else {
    old_text <- clean_arg(old)
  }
  
  tibble::tibble(
    level = level,
    version = clean_arg(version),
    old = old_text,
    new  = new_text
  )
}
parse_script <- function(path) {
  lifecycle_calls <- path |>
    parse(keep.source = TRUE) |>
    xmlparsedata::xml_parse_data(pretty = TRUE) |>
    xml2::read_xml() |>
    xml2::xml_find_all(
    ".//SYMBOL_FUNCTION_CALL[contains(text(), 'deprecate_')]"
  ) 
  
  if (length(lifecycle_calls) == 0) return(NULL)
  
  lifecycle_calls_info <-  purrr::map(lifecycle_calls, parse_lifecycle_call)

  do.call(rbind, lifecycle_calls_info)
}

scripts <- fs::dir_ls(here::here("R"), glob = "*.R")

table <- do.call(
  rbind,
  purrr::map(scripts, parse_script), 
  args = list(make.row.names = FALSE)
)

knitr::table(table)

```
