---
title: "Plotting graphs"
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6,
  fig.align = "center",
  dev = "png"
)
set.seed(42) # stable force-directed layouts across rebuilds
```

```{r setup}
library(igraph)
```

`igraph` ships a native, dependency-light plotting engine built on base R
graphics. A single call to `plot()` "just works", and every visual detail —
vertex shapes, edge routing, colours, labels, legends — is controllable through
arguments to `plot()` or through vertex/edge/graph attributes.

This article is a complete tour of that engine: it walks through both the
long-standing features and the more recent additions (scales and legends,
label repelling and halos, richer edge styles, colour gradients and
transparency, and SVG export). Everything is shown at least once.

```{r}
# A small attributed social network used throughout this article.
g <- make_graph(
  ~ Alice - Boris:Himari:Moshe, Himari - Alice:Nang:Moshe:Samira,
  Ibrahim - Nang:Moshe, Nang - Samira
)
V(g)$gender <- c("f", "m", "f", "m", "m", "f", "m")
E(g)$kind <- rep(c("friend", "colleague"), length.out = ecount(g))
```

## Quick start

The simplest possible call lays the graph out automatically and draws it:

```{r}
plot(g)
```

There are two ways to control appearance:

* **As an argument to `plot()`**, prefixed with `vertex.` / `edge.` /
  (none for plot-level), e.g. `vertex.color = "tomato"`.
* **As a graph attribute**, e.g. `V(g)$color <- "tomato"`.

Arguments to `plot()` take precedence over attributes, which in turn take
precedence over options and built-in defaults (see
[Setting defaults](#setting-defaults)). Keeping the styling in the `plot()`
call keeps the visual representation separate from the graph data:

```{r}
plot(
  g,
  vertex.color = ifelse(V(g)$gender == "m", "skyblue", "pink"),
  vertex.size = 30,
  edge.width = ifelse(E(g)$kind == "colleague", 3, 1)
)
```

## Layouts

A graph is an abstract object; to draw it we first map vertices to coordinates
with a *layout algorithm*. All layout functions start with `layout_`:

| Method name        | Algorithm description                                          |
|--------------------|---------------------------------------------------------------|
| `layout_randomly`  | Places vertices completely randomly                           |
| `layout_in_circle` | Deterministic; places vertices on a circle                    |
| `layout_on_sphere` | Deterministic; places vertices evenly on a sphere surface     |
| `layout_with_drl`  | DrL (Distributed Recursive Layout), for large graphs          |
| `layout_with_fr`   | Fruchterman–Reingold force-directed algorithm                 |
| `layout_with_kk`   | Kamada–Kawai force-directed algorithm                         |
| `layout_with_lgl`  | LGL (Large Graph Layout), for large graphs                    |
| `layout_as_tree`   | Reingold–Tilford tree layout                                  |
| `layout_nicely`    | Picks a sensible algorithm automatically (the default)        |

A layout function returns a matrix with one row per vertex. You can pass either
a precomputed matrix or the function itself:

```{r}
lay <- layout_with_kk(g)
plot(g, layout = lay, main = "Kamada-Kawai layout")
```

```{r}
plot(g, layout = layout_in_circle, main = "Circular layout")
```

By default coordinates are rescaled to fit a `[-1, 1]` square with aspect ratio
1. Disable rescaling and set the limits/aspect/margins yourself when you need
the layout's own coordinates:

```{r}
plot(
  g,
  layout = lay,
  rescale = FALSE,
  xlim = range(lay[, 1]),
  ylim = range(lay[, 2]),
  asp = 1,
  margin = 0.1
)
```

## Vertices

Vertex colour, size and frame are the basic knobs. `vertex.size2` sets the
second dimension for non-square shapes (e.g. rectangles).

```{r}
plot(
  g,
  vertex.color = "gold",
  vertex.size = 35,
  vertex.frame.color = "grey30",
  vertex.frame.width = 2
)
```

### Transparency

`vertex.alpha` (and `edge.alpha`, later) set opacity in `[0, 1]`, folded into
the fill colour:

```{r}
plot(g, vertex.color = "purple", vertex.alpha = 0.4, vertex.size = 40)
```

### Shape gallery

The built-in shapes are `circle`, `square`, `csquare` (centred square),
`rectangle`, `crectangle` (centred rectangle), `vrectangle` (vertical
rectangle), `none` (labels only), `pie`, `sphere`, and `raster`. Use `shapes()`
to list everything currently registered:

```{r}
shapes()
```

```{r fig.width = 7, fig.height = 7}
shp <- c(
  "circle", "square", "csquare", "rectangle",
  "crectangle", "vrectangle", "none", "pie", "sphere", "raster"
)
ring <- make_ring(length(shp))
plot(
  ring,
  vertex.shape = shp,
  vertex.label = shp,
  vertex.label.dist = 1.4,
  vertex.size = 26,
  vertex.size2 = 18,
  # pie data is only used by the "pie" shape
  vertex.pie = lapply(shp, function(s) if (s == "pie") c(2, 3, 1) else 0),
  vertex.pie.color = list(c("tomato", "gold", "skyblue"))
)
```

The `pie` shape draws a small pie chart per vertex; its slices, colours, shading
angle, shading density and line type are controlled by `vertex.pie`,
`vertex.pie.color`, `vertex.pie.angle`, `vertex.pie.density` and
`vertex.pie.lty`. The `raster` shape draws an image (defaulting to the igraph
logo); supply your own with `vertex.raster`.

### Relative sizing

By default `vertex.size` is in absolute units. Set `vertex.size.scaling = TRUE`
to size vertices relative to the plot region, controlled by `relative.size`:

```{r}
plot(
  make_star(8),
  vertex.size.scaling = TRUE,
  vertex.relative.size = c(0.02, 0.06)
)
```

## Vertex labels

Labels default to vertex names (or IDs). They are positioned at a distance
`label.dist` from the vertex, in the direction `label.degree` (an angle in
radians), and styled with the usual font controls. `label.angle` rotates the
text and `label.adj` fine-tunes its anchor.

```{r}
plot(
  g,
  vertex.size = 30,
  vertex.label.color = "black",
  vertex.label.family = "sans",
  vertex.label.font = 2, # bold
  vertex.label.cex = 1.1,
  vertex.label.dist = 2.2,
  vertex.label.degree = -pi / 2 # place labels above the vertices
)
```

`vertex.label.angle` rotates the label text (in degrees) and `vertex.label.adj`
adjusts its anchor, exactly like `srt` and `adj` in `graphics::text()`:

```{r}
plot(
  g,
  vertex.size = 36,
  vertex.label.angle = 30,
  vertex.label.adj = c(0.5, 0.5)
)
```

### Avoiding overlaps with `label.repel`

On dense graphs labels collide. Set `vertex.label.repel = TRUE` to iteratively
nudge overlapping labels apart, drawing a thin leader line from each moved label
back to its vertex:

```{r}
gr <- make_ring(12)
V(gr)$name <- paste0("node-", seq_len(vcount(gr)))
plot(gr, layout = layout_in_circle, vertex.size = 8, vertex.label.repel = TRUE)
```

### Legibility with `label.halo`

`vertex.label.halo` draws a halo (outline) behind the label text so it stays
readable over edges and other vertices. `vertex.label.halo.width` controls the
outline thickness as a fraction of the label height. The default `NA` draws no
halo.

```{r}
plot(
  g,
  vertex.size = 30,
  vertex.color = "steelblue",
  vertex.label.color = "white",
  vertex.label.halo = "black",
  vertex.label.halo.width = 0.3
)
```

## Edges

Edges are styled with colour, width and line type, and (for directed graphs)
arrowheads. `edge.arrow.mode` chooses the arrow direction (`0` none, `1`
backward, `2` forward, `3` both, or the symbolic `"<"`, `">"`, `"<>"`); it also
accepts `"a:attr"` to read the mode from a vertex attribute.

```{r}
dg <- make_graph(~ A -+ B, B -+ C, C -+ A, A -+ D)
plot(
  dg,
  edge.color = "grey40",
  edge.width = 2,
  edge.lty = "dashed",
  edge.arrow.size = 0.8,
  edge.arrow.width = 1.2
)
```

### Curved edges and multi-edge fans

`edge.curved` bends edges: `TRUE` means curvature 0.5, numeric values set the
amount (negative bends the other way). For graphs with parallel edges,
`curve_multiple()` fans them apart so they don't overlap:

```{r}
multi <- graph_from_edgelist(
  matrix(c(1, 2, 1, 2, 1, 2, 2, 3), ncol = 2, byrow = TRUE)
)
plot(multi, edge.curved = curve_multiple(multi), edge.arrow.size = 0.6)
```

### Edge styles

The `edge.style` argument selects an edge-routing style:

* `"auto"` (default) — straight, or an arc when `edge.curved` is non-zero
* `"straight"` — always straight
* `"arc"` — a curved arc whose height comes from `edge.curved`
* `"elbow"` — a two-corner orthogonal (right-angle) connector
* `"diagonal"` — a smooth S-curve with axis-aligned ends

The orthogonal and diagonal styles are well suited to tree- and grid-like
layouts:

```{r fig.width = 7, fig.height = 3.2}
tree <- make_tree(7, children = 2)
op <- par(mfrow = c(1, 3), mar = c(1, 1, 2, 1))
plot(tree, layout = layout_as_tree, edge.style = "arc", main = "arc")
plot(tree, layout = layout_as_tree, edge.style = "elbow", main = "elbow")
plot(tree, layout = layout_as_tree, edge.style = "diagonal", main = "diagonal")
par(op)
```

### Transparency and colour gradients

`edge.alpha` sets edge opacity. `edge.gradient = TRUE` colours each edge with a
gradient running from the source vertex's colour to the target's — a way to show
direction without arrowheads:

```{r}
ring_d <- make_ring(8, directed = TRUE)
V(ring_d)$color <- rainbow(vcount(ring_d))
plot(
  ring_d,
  layout = layout_in_circle,
  vertex.size = 18,
  edge.gradient = TRUE,
  edge.width = 4,
  edge.arrow.mode = 0
)
```

```{r}
plot(dg, edge.color = "navy", edge.alpha = 0.3, edge.width = 6)
```

### Self-loops

Loops are drawn as small arcs. `edge.loop.angle` sets the direction of a loop
(in radians) and the plot-level `loop.size` scales them:

```{r}
loops <- make_graph(~ A -+ A, A -+ B, B -+ B)
plot(loops, edge.loop.angle = pi / 4, loop.size = 1.2, edge.arrow.size = 0.6)
```

## Edge labels

Edge labels are added with `edge.label` and styled with the matching
`edge.label.*` controls, including a halo (`edge.label.halo`) just like vertex
labels. `edge.label.x` / `edge.label.y` override the automatic placement.

```{r}
plot(
  dg,
  edge.label = c("a", "b", "c", "d"),
  edge.label.color = "black",
  edge.label.cex = 1.2,
  edge.label.family = "sans",
  edge.label.halo = "yellow",
  edge.label.halo.width = 0.4,
  edge.arrow.size = 0.6
)
```

## Colours and palettes

igraph provides perceptually sensible palettes:
`categorical_pal()` for unordered categories, `sequential_pal()` for ordered
magnitudes, `diverging_pal()` for signed data, and `r_pal()` for R's classic
palette.

```{r fig.width = 8, fig.height = 2.4}
op <- par(mfrow = c(1, 4), mar = c(0, 1, 2, 1))
pie(rep(1, 8), col = categorical_pal(8), main = "categorical_pal(8)")
pie(rep(1, 9), col = sequential_pal(9), main = "sequential_pal(9)")
pie(rep(1, 9), col = diverging_pal(9), main = "diverging_pal(9)")
pie(rep(1, 8), col = r_pal(8), main = "r_pal(8)")
par(op)
```

When a vertex/edge colour is given as a factor or integer, it indexes into the
active `palette`:

```{r}
plot(
  g,
  vertex.color = as.factor(V(g)$gender),
  palette = categorical_pal(8),
  vertex.size = 30
)
```

## Scales and legends

Passing a raw colour vector means igraph never learns what the colours *mean*,
so it cannot explain them. The `scale_*()` helpers fix this: they map a data
column to an aesthetic **and record the mapping**, so `plot()` can draw a
matching guide automatically.

`scale_color()` on a discrete variable produces a categorical legend:

```{r}
plot(
  g,
  vertex.color = scale_color(V(g)$gender, name = "Gender"),
  vertex.size = 30
)
```

On a numeric variable it produces a continuous **colourbar**:

```{r}
plot(
  g,
  vertex.color = scale_color(degree(g), name = "Degree"),
  vertex.size = 30
)
```

`scale_size()` maps a numeric variable to a size range and draws a size legend:

```{r}
plot(
  g,
  vertex.size = scale_size(degree(g), range = c(10, 35), name = "Degree"),
  vertex.color = "lightgrey"
)
```

The `legend` argument controls placement: `TRUE` (default, right side),
`"left"`, `"top"`, `"bottom"`, or `FALSE` to suppress the guide. Multiple scales
combine into one guide region:

```{r}
plot(
  g,
  vertex.color = scale_color(V(g)$gender, name = "Gender"),
  vertex.size = scale_size(degree(g), range = c(12, 34), name = "Degree"),
  legend = "bottom"
)
```

## Decluttering with `label_top()`

On large graphs, labelling every vertex is unreadable. `label_top()` keeps only
the highest-ranking labels by some metric and blanks the rest with `NA` (which
`plot()` omits):

```{r}
big <- sample_gnp(40, 0.08)
plot(
  big,
  vertex.size = 6,
  vertex.label = label_top(degree(big), n = 5),
  vertex.label.dist = 1,
  vertex.label.color = "black"
)
```

Use `prop` to keep a fraction instead of a count, and `decreasing = FALSE` to
keep the lowest:

```{r}
plot(
  big,
  vertex.size = 6,
  vertex.label = label_top(degree(big), prop = 0.1, decreasing = FALSE),
  vertex.label.dist = 1,
  vertex.label.color = "black"
)
```

To label everything above a fixed cutoff you don't need the helper —
`ifelse(metric > cutoff, labels, NA)` works directly.

## Highlighting groups

`mark.groups` draws a smoothed, filled polygon "under" a set of vertices. The
fill, border, smoothness, padding and border width are set by `mark.col`,
`mark.border`, `mark.shape`, `mark.expand` and `mark.lwd`:

```{r}
plot(
  g,
  vertex.size = 25,
  mark.groups = list(c("Alice", "Himari", "Moshe"), c("Nang", "Samira")),
  mark.col = c("#ffd9d9", "#d9ecff"),
  mark.border = c("tomato", "steelblue"),
  mark.shape = 1 / 2,
  mark.expand = 20,
  mark.lwd = 2
)
```

## Annotations and frame

Plot-level titles and axes follow base graphics conventions: `main`, `sub`,
`xlab`, `ylab`, `axes` and `frame.plot`.

```{r}
plot(
  g,
  vertex.size = 25,
  main = "Friendship network",
  sub = "an igraph example",
  xlab = "x",
  ylab = "y",
  axes = TRUE,
  frame.plot = TRUE
)
```

## Custom vertex shapes

You can register your own shape with `add_shape()`. A shape is defined by a
*clip* function (where edges stop) and a *plot* function (how the vertex is
drawn); `shape_noclip` and `shape_noplot` are no-op building blocks, and
`shapes("circle")$clip` reuses an existing clipper.

```{r}
mytriangle <- function(coords, v = NULL, params) {
  vertex.color <- params("vertex", "color")
  if (length(vertex.color) != 1 && !is.null(v)) {
    vertex.color <- vertex.color[v]
  }
  vertex.size <- params("vertex", "size")
  if (length(vertex.size) != 1 && !is.null(v)) {
    vertex.size <- vertex.size[v]
  }
  symbols(
    x = coords[, 1], y = coords[, 2], bg = vertex.color,
    stars = cbind(vertex.size, vertex.size, vertex.size),
    add = TRUE, inches = FALSE
  )
}

add_shape("triangle", clip = shapes("circle")$clip, plot = mytriangle)

plot(
  g,
  vertex.shape = "triangle",
  vertex.color = "seagreen",
  vertex.size = 40
)
```

`shape_noclip` (edges run all the way to the centre) and `shape_noplot` (draw
nothing) are the no-op building blocks; registering a shape from them gives an
"invisible" vertex that still carries a label:

```{r}
add_shape("blank", clip = shape_noclip, plot = shape_noplot)
plot(g, vertex.shape = "blank", vertex.label.dist = 0)
```

## Setting defaults {#setting-defaults}

`igraph_options()` sets defaults for every subsequent plot, and `igraph_opt()`
reads one back. Plotting options use the same `vertex.` / `edge.` prefixes:

```{r}
old <- igraph_options(
  vertex.color = "orange",
  vertex.size = 30,
  edge.color = "grey70"
)
plot(g)
igraph_opt("vertex.color") # read one option back
igraph_options(old) # restore
```

The full precedence order, from highest to lowest, is:

1. an explicit argument to `plot()`,
2. a vertex/edge/graph attribute,
3. an `igraph_options()` setting,
4. the built-in default.

Setting `igraph_options(annotate.plot = TRUE)` makes `plot()` annotate the
figure with the graph's name and its vertex/edge counts by default.

## SVG export with tooltips

`as_svg()` renders a graph to a standalone SVG. Beyond crisp vector output, it
tags each vertex and edge with element IDs and a `<title>`, so the SVG has
hover tooltips with **no JavaScript dependency**. The `tooltips` argument picks
which vertex attribute to use for the titles (defaulting to the vertex name).

```{r}
svg <- as_svg(g, width = 6, height = 6, tooltips = "gender")
# write to a file ...
tmp <- tempfile(fileext = ".svg")
cat(svg, file = tmp)
substr(svg, 1, 60)
```

Each vertex becomes `<g id="vertex-N"><title>...</title>...</g>` and each edge a
matching group, so downstream tools (or a browser) can offer hover and click
interactions. Embedded directly into this page, hovering a vertex shows its
tooltip:

```{r results = "asis", echo = TRUE}
cat(as_svg(g, width = 6, height = 6, tooltips = "gender", vertex.size = 30))
```

## Where to go next

* The per-function reference: <https://r.igraph.org/reference/>, in particular
  `?plot.igraph` and `?igraph.plotting` for the full list of parameters.
* The main [igraph](igraph.html) introduction for graph construction and
  analysis.

```{r}
sessionInfo()
```
